### 2019-12-12 review_article.Rmd
editing to camparisons of alternative displays

---
title: "Cancer Applications of Choropleth maps, and the Potential of Cartograms and Alternative Map Displays"
author: "Stephanie Kobakian, Jessie Roberts and Dianne Cook"
bibliography: reviewbib.bib
output:
  pdf_document:
    fig_caption: yes
    number_sections: yes
    toc: no
    toc_depth: 3
  word_document:
    toc: no
    toc_depth: '3'
link-citations: yes
fontsize: 11pt
csl: springer-vancouver.csl
yolo: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
library(cowplot)
library(png)
library(grid)
```

\section*{Abstract}
<!-- 105 words -->

Cancer atlases communicate cancer statistics over geographic domains. They subdivide these domains into administrative regions such as countries, states, or suburbs. When communicating human-related statistics in Australia, the geographic map base draws attention to sparsely populated rural areas. The smaller geographic areas may not draw the attention of readers, but they are important to consider if they are more densely populated. Alternative map displays can be more effective than a choropleth map at revealing spatial patterns. Alternative displays are effective for the communication of election data aggregated at the state or electorate level on online news sites.

# Introduction
<!-- 436 words -->
<!-- What data are we working with, why, what are the purposes -->
Researchers, health authorities, governments and not-for-profits are common communicators of cancer statistics. They often present statistics to the public as aggregated values for geopolitical areas. Presenting these statistics requires aggregating individual observations into the statistics for the geographical units, especially for privacy protection, but also for political and policy purposes.
Data collection or aggregation regions for the amount of cancer diagnoses or excess deaths per area include states, provinces, local government areas, and post/zip codes. To provide the amount, or count, of the diagnoses or excess deaths with these areas, it needs minimal transformation. However, the amount of people living in each of the areas in a set will be different. Those who wish to calculate rates can use the amount of people living in an area as a denominator and the diagnoses can become a numerator to create rates per amount of people. Health registries routinely collect data for public health reasons and may make this available to the public as a service to the community.

<!-- Ways to visualize, whats our hypothesis -->
Choropleth map displays are commonly used for the communication of cancer statistics over geographic domains. In this display, map creators map a statistic to color and apply the colour to the polygons representing geographic regions. Map users may overlook geographic patterns of disease when obscured in a table [@SAMGIS]. Visualizing diseases on maps is often the first step in exploratory spatial data analysis and helps in the formulation of hypotheses. Providing a visual representation of cancer outcomes allows identification, and public health policy and actions can address these patterns. choropleth maps allow the examination of the spatial distribution of the disease incidence, this may reveal a trend in longitude or latitude, differences between rural areas and urban, or coastal and inland areas, or even specific hot spots of the disease. Daniel Exeter [@SE] recognizes that one of the key challenges with mapping spatial patterns of disease is the design of visualizations. This paper considers the current visualization techniques to communicate statistics to members of the public and their applications to cancer statistics. It also considers alternative approaches and compares the visualization methods, highlighting the differences and historic use of these displays.

<!-- Structure of the paper -->
The paper is structured as follows. The next section describes the choropleth map, which is the common approach to disease maps and presents examples of atlases in use today. Section 3 discusses the limitations of the choropleth map. Section 4 describes alternative displays, including the cartogram, which is useful when the map has heterogeneously sized geographic units. Section 5 presents the limitations of the production and use of alternative displays. Disease maps are more useful when made interactive, Section 6 describes common options, along with a discussion of benefits and disadvantages.

# Traditional approaches for cancer map displays {#ch:choropleth maps}
<!-- 381 words -->
<!-- what is a choropleth map -->
A choropleth is traditionally used to make a cancer map. It is used to show the geographical distribution of data values by shading areas of a map [@EI]. A choropleth map involves faithfully rendering the geography, and the design of the color scheme and rendering reveal spatial patterns among data values. Identifying and explaining spatial structures, patterns, and processes involve considering the individuals and organizing them into representable units of communities [@SAMGIS]. Early versions of choropleth maps used symbols or patterns instead of color. choropleth maps can be used for displaying disease data [@DMAHP], including cancer data [@CPISACA]. In epidemiology, choropleth maps are often used as a tool to study the spatial distribution of cancer incidence and mortality.

Figure 2<!--\ref{fig:choroCRS}--> shows various geographic projections of choropleth maps of age-adjusted rate (per 100,000 people) of new cases of lung and bronchus in the USA, averaged from 2012 to 2016. The official federal statistics provide cancer incidence and deaths [5], produced by the Centers for Disease Control and Prevention (CDC) and the National Cancer Institute (NCI). The rates exhibit a spatial trend, increasing from west to east. Utah is a spatial outlier, it has a noticeably lower rate than its neighbors. Also, Kentucky and Maine have higher rates than their neighbors. There is possibly a cluster of higher rates around the tobacco states.

<!-- psychology -->
Displaying familiar state boundaries can make a map easier to read [@CIBMUK] and allow viewers to infer the spatial relationships visually in the data using their mental model of the geography.
Consumers of disease displays may include researchers, members of the public, policymakers, and the media  [@CPISACA]. For these consumers, the familiarity of the geography is a worthy consideration when presenting results of spatial analysis.

<!-- potential problems -->
Politically driven boundaries produced the now familiar shapes or regions, this has created individual areas of non-uniform size. These areas have different population densities and are subject to change. 
The different population and geographical sizes of administrative areas can attract attention to the shades of the underpopulated but large areas [@EI], Skowronnek [@BCM] calls this an area-size bias. 
choropleth maps can inhibit visual inference when presenting human-related statistics as the display may draw attention away from the 'potentially more important results in the more populous communities' that are geographically smaller [@SE].

```{r choropleth-grid-create, message=FALSE, warning = FALSE, fig.cap = "Examples of publicly available choropleth cancer maps.", out.width='100%', fig.align='center', fig.width=12, fig.height=10}
library(cowplot)

# Import the images
plot1 <- png::readPNG('cancer_map_images/choropleths/21.choropleth.png')
plot2 <- png::readPNG('cancer_map_images/choropleths/3.choropleth.png')
plot3 <- png::readPNG('cancer_map_images/choropleths/25.choropleth.png')
plot4 <- png::readPNG('cancer_map_images/choropleths/15.choropleth.png')
plot5 <- png::readPNG('cancer_map_images/choropleths/7.choropleth.png')
plot6 <- png::readPNG('cancer_map_images/choropleths/23.choropleth.png')
plot7 <- png::readPNG('cancer_map_images/choropleths/26.choropleth.png')

ggdraw() +
  draw_plot(rasterGrob(plot1), 0, .75, 0.33, .25) +
  draw_plot(rasterGrob(plot2), 0.33, 0.75, 0.33, .25) +
  draw_plot(rasterGrob(plot3), 0.66, 0.5, .33, .5) +
  draw_plot(rasterGrob(plot4), 0, 0.5, .33, .25) +
  draw_plot(rasterGrob(plot5), 0.33, 0.5, .33, .25) +
  draw_plot(rasterGrob(plot6), 0, 0, .66, .5) +
  draw_plot(rasterGrob(plot7), .66, 0, .33, .5) +
  draw_plot_label(c("a", "b", "c", "d", "e", "f", "g"), 
    c(0.015, 0.33, 0.63, 0.015, 0.33, 0.014, 0.63), 
    c(0.99, 0.99, 0.99, 0.74, 0.74, 0.47, 0.47), size = 20)

ggsave(filename = "figures/choropleth_grid.png", 
  device = "png", width = 6, height = 5, dpi = 300)
```

## Cancer atlases {#ch:public}
<!-- 360 words -->
<!-- Displaying of cancer data on choropleth maps -->
A cancer atlas is a map, or collection of maps, representing cancer incidence and mortality for a country or group of countries. Atlases are key to developing hypotheses regarding areas with unusually high rates, geographic correlations, work-related exposures, and high-risk diets [@MACM]. The data collection methods across regions and the administrative control within regions lends itself to choropleth visualization. 
Cancer maps and atlases dating back to Haviland's maps in 1875, and early work in US cancer atlases appearing in 1971 [@burbank]. The presentation of cancer statistics has increased with greater access to computational power and the availability of geographic information systems software [@SE].

<!-- Measures reported in cancer atlases -->
Cancer maps are effective tools for communicating incidence, survival, and mortality to a wide range of audiences, including the general public and others not trained in statistical analysis.
These visualizations enable non-expert audiences to interpret the outputs of sophisticated statistical analysis. Cruickshank (1947) as cited by S. D. Walter [@DMAHP], discusses using visuals as a 'formal statistical assessment of the spatial pattern'. Overwhelmingly, the cancer maps used to communicate to members of the public and other non-expert audiences are choropleth maps.

Epidemiologists and statisticians have developed these statistics over several decades. Mortality rates commonly presented as relative rates of risk across the population and age-adjusted to correct for the higher prevalence of cancers in older populations.
Howe [@HEDP] describes Stock's development of the standardized mortality ratios through the 1930s. Table 1<!--\ref{tab:measures}--> summarizes the measures commonly presented in published cancer atlases.

Table: 1<!--\label{tab:measures}--> Common measures for reporting cancer information.

|Measure |Details| 
|:--------------------|:------------------------------------------------|  
|1. Count| Crude cancer counts|
|2. Rate per 100,000| Cancer incidence per 100,000 population|  
|3. IR (Incidence Ratio)| $(IR)_i=\frac{(Incidence\ Rate)_i}{Average\ Incidence\ Rate}$, |
|| The cancer incidence rate in region $i$ over the average cancer incidence rate for all of the regions|
|4. Age-Adjusted Rate per 100,000| Standardized by age structure or region|   
|5. Age-Adjusted Relative Risk| Standardized by age structure in each region $i$|  
|6. SIR (Standardized Incidence Ratio)| Standardized by incidence in each region $i$|
|7. Below or above Expected| An alternative expression of the SIR|
|8. RER | $RER = \frac{(Cancer\ related\ mortality)_i}{Average\ cancer\ related\ mortality}$|
|(Relative Excess Risk)| Represents the estimate of cancer-related mortality within five years of diagnosis. Also referred to as 'excess hazard ratio'|

<!-- 416 words -->
Roberts [@roberts2019communication] identified 33 publicly available cancer atlases, published between 2010 and 2018, from different countries around the globe. These online atlases used choropleth maps. Non-commercial organizations, including not-for-profits, government, research organizations, advocacy groups or government-funded partnerships published all but one of these. Figure 1<!--\ref{fig:choropleth-grid-create}--> displays a subset of maps from these atlases.

The selection varies in the geographies explored, Figure 1<!--\ref{fig:choropleth-grid-create}-->b, Globocan 2012 [@Globocan] explores Estimated Cancer Incidence, Mortality and Prevalence Worldwide utilizing data sourced from cancer registries of each country. 
The Bowel Cancer Australia Atlas presents an example of a cancer specific atlas, Figure 1<!--\ref{fig:choropleth-grid-create}-->d shows the average Standardized Incidence Ratio of colorectal cancer for Australian males from 2006 to 2010 in Australia [@Bowel]. Like many of the atlases examined, there is a choice of gender displayed in the Bowel Cancer Atlas. An alternative approach used in the Map of Cancer Mortality Rates in Spain, Figure 1<!--\ref{fig:choropleth-grid-create}-->f shows side-by-side maps of the relative risk of lung cancer separately for men and women [@cancerSpain].

There are vast differences in the geographies considered by online atlases. 
Figure 1<!--\ref{fig:choropleth-grid-create}-->b shows global age-standardized incidence rates (per 100,000) for all invasive cancers for both men and women, aggregated at a national level for 2018.
The United States Cancer Statistics [@USInteractive] presented in Figure 1<!--\ref{fig:choropleth-grid-create}-->e contains the incidence rate per 100,000, of all cancer types for men and women in the United States in 2016, aggregated at the state level.
The Environment and Health Atlas of England and Wales, Figure 1<!--\ref{fig:choropleth-grid-create}-->a shows the relative risk for women developing lung cancer in England and Wales in 2010 [@EnvEnglandWales2] at a neighborhood (small-area) scale.
The Atlas of Cancer in Queensland, Figure 1<!--\ref{fig:choropleth-grid-create}-->c shows the relative incidence ratio of lung cancer in males for each Statistical Area at Level 2 in the state of Queensland within Australia [@QLDcancerAtlas].

Age-specific atlases are less common, Figure 1<!--\ref{fig:choropleth-grid-create}-->g displays Atlas of Childhood Cancer in Ontario, this communicates the incidence rate of childhood cancers per 100,000 (by census division) for children aged 0-14, in Ontario from 1995 to 2004 [@OntarioPediatric]. This atlas communicates information regarding a tiny geographic region.

## Additional materials
<!-- 479 words -->
Including supplementary graphs and plots add more depth and information. Additional materials such as tables, graphs, and text explanations, support the understanding and inference derived from maps, ensuring the message communicated will be consistent across a range of viewers [@CPISACA]. The many displays of statistical summaries, including dot plots, bar plots, box plots, cumulative distribution plots, scatter plots, normal probability plots, can provide alternative views of the cancer statistics. These can also display supporting statistics such as error, confidence intervals, distributions, sample or population sizes, standard deviation. When presenting cancer maps the display of the supplementary statistics helps. The intuition derived from maps must be ‘validated by rigorous statistical analyses’ [@MACM].

The statistics communicated in atlases are often used to describe differences between areas. This can occur at different levels of aggregation. Aggregation of global health statistics occurs within administrative and arbitrarily defined regions, such as those used by the World Health Organization and the United Nations [@IARC_3].
World atlases can allow for displays of data aggregated into continents, countries, states, provinces and congressional districts [@USInteractive].
Each population area will probably have a different number of people. Cancer atlases may also communicate the distribution of the population living in all areas in a table or histogram display [@NICR_1].
Atlases can connect the population to the land available to them by communicating population density. Section  4<!--\ref{ch:alternatives}--> considers alternatives to choropleth that introduce population into the visualizations.

Map users can define the population further using the demographics of the individuals within each area.  Demographics that can influence cancer rates include information regarding the age and sex distribution of the areas presented. Some digital atlases surveyed allow users to control the display to select subsets such as males, females or those aged over 65. 
Similarly, socioeconomic indicators, such as unemployment rates, poverty rates, remoteness, and education levels,  can explain how the experience of cancer prevalence varies for various members of society. Few atlases provide this level of detail.
Introducing population and demographic information helps to interpret the rates in areas effectively, but there will still be uncertainty around the rates.
To address this, cancer atlases often communicate uncertainty about the value of a statistic. Cancer atlas creators may introduce uncertainty when using samples, or there are errors occurring in the disease reporting, or models are used to simulate values to maintain privacy. The statistics that report on uncertainty include sample or population size, standard deviation, confidence intervals, statistical significance, quantiles, and  credible intervals. The most common measures used to represent uncertainty were credible or confidence intervals (CIs). Displaying the uncertainty associated with reported statistics is a vital feature of a cancer map, but it is difficult to display effectively. The map focuses on displaying the statistic and lacks additional space to represent the uncertainty. Providing an adjacent map or overlaying maps with symbols [@VSSDCUC] are two approaches.

# Limitations of choropleth displays {#ch:chorolimit}

The administrative boundaries used to define regions may limit a choropleth display.
Those who critique choropleth maps often dicuss election mapping, where each electorate decides the outcome of one parliamentary seat in the house of representatives.
While all states are visible in a map display of Australia, smaller regions such as electorates present an urban-rural divide, where rural regions are much larger than metropolitan areas of similar or greater populations. 
The landmass occupied by urban electoral districts is only 10% of Australia, yet 90% of the population live in these urban areas. To present election results on a choropleth map should be 'unthinkable',XXX as it means diminishing the visual impact of most of the electorates.
Using a map of Australia as a visual statistic would communicate an incorrect distribution of votes across the country.

Spatially heterogeneous population distribution across a country's landmass is an issue common to many countries, especially British colonies [@ACTUC]. However, the boundaries shown in choropleth maps provide a familiar display. This may be less threatening to an audience that could be unfamiliar was interpreting spatial statistics, even though they suffer primarily from unfaithfully representing the disease information by hiding small geographic areas. Inset maps like in Brisbane city in Figure 1<!--\ref{fig:choropleth-grid-create}-->c of the state of Queensland are commonly used to reduce distorted interpretations, but it is a band aid remedy. 
As this feature of population distributions continues to intensify, the need for cartograms as an alternative to a choropleth map should only increase.

# Contemporary alternatives to choropleth maps {#ch:alternatives}

## Cartograms
<!-- 1496 words inc table-->

<!-- What is a cartogram? -->
choropleth maps imply uniformity of data across the geographic space but population densities are unlikely to be uniform [@BCM]. Cartographers developed the cartogram to draw the attention to the population by transforming the map [@ACCAC]. The resulting display can communicate the impact of the disease more accurately across the population, as recorded by the statistic, at the sacrifice of geographic accuracy.

<!-- There are many alternative views of the world-->
<!-- Tobler map projections - I think this is not relevant here
choropleth maps may be considered true topological maps however,, if the landmass displayed covers enough of the globe, there must be a transformation or distortion to display the land in 2D [@HTLWM]. The amount of distortion is related to the distance covered by the landmass displayed @GAMP.
World map projections reflect the frequent perspectives used to view the earth. Choropleth maps will always be distorted if they cover enough of the globe, just like photographs of the globe from space. Choropleth creation requires choosing a map projection that shows a favorable distortion of the geography for presenting the set of spatial information. Selecting a display can prevent misinterpretation of global statistics, as global maps face the challenge of equitable displays of landmass on maps [@RSCW].
If the statistic presented on the map base relies on physical distance and is influenced by the topology there is no transformation needed, beyond choosing a reasonable projection. 
-->

```{r choroCRS, fig.cap = "Four choropleth maps average age-adjusted rate of incidence for lung and bronchus in the United States averaged from sex and years 2012-2016, using four different map projections. The map projections alter the shapes and angles of the boundaries of each state.", message=FALSE, warning=FALSE, out.width='100%', fig.align='center'}
#knitr::include_graphics("figures/ggchoroCRS.png")
c1 <- png::readPNG("figures/gg3857.png")
c2 <- png::readPNG("figures/gg2163.png")
c3 <- png::readPNG("figures/gg4326.png")
c4 <- png::readPNG("figures/gg2955.png")
#knitr::include_graphics("figures/usa_grid.png")
load("figures/usa_legend.rda")

ggdraw() +
  draw_plot(rasterGrob(c1), 0, .55, 0.5,    0.45) +
  draw_plot(rasterGrob(c2), 0.5, 0.55, 0.5, 0.45) +
  draw_plot(rasterGrob(c3), 0.0, 0.1, 0.5, 0.45) +
  draw_plot(rasterGrob(c4), 0.5, 0.1, 0.5,   0.45) +
  draw_plot(usa_legend, 0, 0, 1, 0.1, scale = 2) +
  draw_plot_label(c("a", "b", "c", "d"), 
    c(0, .5, 0, 0.5), 
    c(1, 1, .5, .5), size = 20)
```

<!-- Denominators -->
When a map creator desires a uniform population density of the map base, the purposeful distortion of the map space is beneficial. The "population distribution is often extremely uneven in former British colonies" [@ACTUC], and this makes the distortion necessary, [@CTTMB] population will then become a uniformly distributed background for the statistic presented [@GOINO].
An area cartogram [@NAC], or population-by-area cartogram [@TAAM] occurs after the distortion of the geographical shape according to population. Event cartograms change the area of regions on a map depending on the amount of disease-related events, but this does not consider the effects of land area and population [@VSSDCUC].

<!-- Why transform?-->
<!-- Common variables used to create cartograms, e.g. population, mortality -->
Cartograms provide an alternative visualization method for statistical and geographical information.
The key difference between a choropleth and a cartogram is the desirable augmentation of the size, shape, or distance of geographical areas [@ACTUC]. 
Monmonier [@HTLWM] suggests that map creators can use white lies to create useful displays by distorting the geometry and suppressing features. It is easy for the average person to disregard the impact of transformations used to create cartograms. One disadvantage of the conventional map is interpreting spatial patterns is difficult because of the sparsely populated rural areas being emphasized in a choropleth map, and the inner city areas are geographically small.
The distortion of each area when making a cartogram is proportional to a value other than the actual earth size area [@NAC]. The distortion in an area cartogram accounts for the population density, preventing it from obscuring the spatial patterns [@TAAM]. 
The spatial transformation of map regions relative to the data emphasizes the data distribution instead of land size [@CBATCC]. When visualizing population statistics, Dorling [@ACTUC] considers this design 'more socially just', or honest [@NISCC], giving equitable representation and attention to all members of the population and reducing the visual impact of large areas with small populations [@DMAHP]. 
Howe [@HEDP] suggests that 'cancer occurs in people, not in geographical areas' and that spatial socio-economic data, like cancer rates, are best presented on a cartogram for urban areas as the population map base avoids allocating 'undue prominence' to rural areas [@CTTMB].

<!-- Overview of varieties -->
<!-- Cartogram makers -->

The creation of cartograms was historically in the hands of professional cartographers [@CD]. 
Early approaches by John Hunter and Jonathan Young (1968) and Durham's wooden tile method, Skoda and Robertson's (1972) steel ball-bearing approach and Tobler's (1973) computer programs [@ACTUC]. 
Howe [@HEDP] discusses the impact of electronic computer-assisted techniques.
Geographical information systems allow map creators to produce cartograms and they use these systems depending on ‘the effectiveness, efficiency, and satisfaction of the map products' (Nielsen 1994) [@CD].

There are many issues to consider when creating alternative map displays: the intended audience of the map, and its purpose are key points in cartogram use and creation as 'there is no "best" cartogram or method of creating cartograms just as there is no "best" map' (Monmonier and Schnell, 1988) [@ACTUC]. 
Nusrat and Kobourov [@SAIC] provided a framework to investigate implementations of the many algorithms presented, and the "statistical accuracy, geographical accuracy, and topological accuracy".

```{r ggcartograms, fig.cap = "Common alternatives to maps, showing the same information as in Figure \ref{choroCRS} for the United States of America: (a) contiguous cartogram, (b) non-contiguous, shape-preserved cartogram, (c)  Dorling cartogram (non-contiguous), (d) hexagon tilemap (non-contiguous). Maps (a) - (c) are created by resizing and reshaping the states of the USA to match the 2015 population of the state. This provides a better sense of the extent of disease relative to the population in the country and can help ease losing information about physically small but population-dense states. Map creators give each state equal size and thus equal emphasis in (d) the hexagon tile map.", message=FALSE, warning=FALSE, out.width='100%'}
g1 <- png::readPNG("figures/ggcont.png")
g2 <- png::readPNG("figures/ggncont.png")
g3 <- png::readPNG("figures/ggdorl.png")
g4 <- png::readPNG("figures/gghexmap.png")
#knitr::include_graphics("figures/usa_grid.png")

load("figures/usa_legend.rda")

ggdraw() +
  draw_plot(rasterGrob(g1), 0, .55, 0.5,  0.45) +
  draw_plot(rasterGrob(g2), 0.5, 0.55, 0.5, 0.45) +
  draw_plot(rasterGrob(g3), 0.0, 0.1, 0.5, 0.45) +
  draw_plot(rasterGrob(g4), 0.5, 0.1, 0.5, 0.45) +
  draw_plot(usa_legend, 0, 0, 1, 0.1, scale = 2) +
  draw_plot_label(c("a", "b", "c", "d"), 
    c(0, .5, 0, 0.5), 
    c(1, 1, .5, .5), size = 20)
```

Table: 2<!--\label{tab:usa}--> Maps used to present statistics for the United States of America. The colour of each state communicates the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.

|Map display |Details| 
|:--------------------|:------------------------------------------------|  
|a. Contiguous| It has distorted each state's shape according to the population of the state in 2015. The state of California has become much larger because of the large population density. This draws attention to the densely populated North-East region and detracts from the less populated Mid West.|
|b. Non-contiguous| It maintains the geographic shape of the states, but the size has altered according to the population of the state in 2015. The state of California has remained closer to its original size than its surrounding states. The North-East states have remained closer to their geographical size, for Massachusetts and Connecticut. This draws attention to the densely populated North-East region and the sparse Mid West.|
|c. Dorling| Circles are used to represent each state, but the population of the state determines the size in 2015.  The North-East states remain closer to their neighbors and are slghtly displaced from their geographic location. It highlights the sparsity of the population in the Mid West by the distance between the circles at the geographic centroids.|
|d. Hexagon Tessellation| A hexagon of equal size represents each state. It is easy to contrast the neighboring states however the North-East regions have been displaced from their geographic location. It highlights the sparsity of the population in the Mid West by the light yellow color, the Age-Adjusted rate in Kentucky is the darkest and its neighbors are similar.|

Figure 3<!--\ref{fig:ggcartograms}--> shows four different cartograms for the same data displayed in Figure 2<!--\ref{fig:choroCRS}-->. The information in Table 2<!--\ref{tab:usa}--> summarizes what can be observed in the four types of cartograms.

### Contiguous

<!-- What is a contiguous cartogram -->
<!-- Intentionally preserve neighbors -->

A contiguous cartogram alters the choropleth according to a statistic and maintains connectivity of the map regions. This transformation often occurs at the expense of the shape of areas [@CBATCC, @NAC, @TAAM].
Min Ouyang and Revesz [@ACA] present three algorithms for creating value-by-area cartograms. They implement 'map deformation' to account for the value assigned to each area.
Other methods include Tobler’s Pseudo-Cartogram Method, Dorling’s Cellular Automaton Method [@ACTUC], Radial Expansion Method of Selvin et al., Rubber Sheet Method of Dougenik et al., Gusein-Zade and Tikunov’s Line Integral Method, Constraint-Based Method (Kocmoud and House) [@CBATCC].

<!-- 
An intentional goal when creating the 1966 Census population cartogram for Canada was to maintain contiguity, while attempting to keep the actual shape of places. The end result was a 'very accurate isodemo-graphic map of Canada'. This intentional design goal coincided with the rising interest in urban geography and presentation of social statistics. Figure 3<!--\ref{fig:ggcartograms}--> a shows a population contiguous cartogram of the United States. All states are visible and the shape of the United States overall is still recognizable.
In contrast, Figure \ref{fig:auscartograms} a shows an Australian contiguous cartogram also based on population. The south east is completely distorted, areas with low population are still large very the map, as their initial area was used to determine the area allocated in the contiguous cartogram.

To be able to recognize the significant changes, a reader will usually have to know the initial geography to find the differences in the new cartogram layout [@NAC]. Tobler's Conformal mapping means to preserve angles locally so that the shapes of small areas on a traditional map and a cartogram would be similar.
@CBATCC presents this issue as conflicting tasks or aims, to adjust region sizes and retain region shapes. 
-->

### Non-contiguous

Non-contiguous cartograms prioritize the shapes of the areas instead of connectivity. Each area stays in a similar position to its location on a choropleth map. Displaying the choropleth map base allows map users to make comparisons regarding the change in the area. The addition is the gap between areas, created as each area shrinks or grows according to the associated value of the statistic. Olson [@NAC] discusses the creation of these maps, the significance of the empty areas left between the geographic boundaries and the new shape, and the 'difference from the original map that is the real message' of these displays.

The distortion of region shapes on the contiguous cartogram presents an additional hurdle to visual recognition, and the non-contiguous cartogram eliminates this hurdle, and presents the meaningful empty-space property [@ECGC, @NAC]. 
<!-- The shapes are valuable for recognition and allows users to orient themselves on the display. Map creators can efficiently communicate with this kind of map by keeping the outlines or particular elements of the original in the new shape [@NISCC].
The scale of the areas does not impact on the shape recognition.
However it may impact on the visibility of all areas if small areas expand beyond their boundaries. 
Figure 3<!--\ref{fig:ggcartograms}-->b shows small overlaps in the North-East region of the United States. These overlaps are also shown in the non-contiguous cartogram of Australia in Figure \ref{fig:auscartograms} b.
-->

### Dorling

Daniel Dorling presents an alternative display engineered to highlight the spatial distribution and neighborhood relationships without complex distortions of borders and boundaries. Daniel Dorling puts forward a simple question and his approach opposes preserving the intricate shape details [@ACTUC]:

>"If, for instance, it is desirable that areas on a map have boundaries which are as simple as possible, why not draw the areas as simple shapes in the first place?"

He acknowledged the sophistication of contiguous cartograms but critiqued their 'very complex shapes,' he answers this with his implementation of maps created using 'the simplest of all shapes'. 
Circular cartograms use the same simple shapes for every region represented and resize the shapes according to the statistic represented or the population for a base map. This familiar shape may be more effective for understanding the spatial distribution than contiguous cartograms. Contiguous cartograms use a gravity model to apply distortions to create the 'nonsense' shapes that have 'no meaning' [@NISCC]. This method applies a gravity model to produce a compelling map, that avoids overlaps and keep spatial relationships with neighboring areas over many iterations. This implementation can work for up to ‘one hundred thousand’ areas.

Raisz laid the groundwork for this approach in the mid-1930s, and his rectangular cartograms provide dramatic comparisons and are especially useful for correcting misconceptions communicated by geographic maps. 
@TFYCC quotes the official definition of Value-Area Cartograms, the simplistic displays which represent each area as a single rectangle, sized according to the value of the statistic.
This rectangular display may sacrifice contiguity but allows for tiling where geographic neighbors placed in suitable relative positions also share borders [@CDWCS]. Rectangular cartograms communicate bi-variate displays of the population by the size of each rectangular, and they use color to communicate a second variable [@ORC].

A similar method to a rectangular cartogram, represents each geographic area using a square. The squares are tessellated to create a grid. This method has been used broadly in the media, e.g. FiveThirtyEight, Bloomberg Business, The Guardian, The Washington Post, The New York Times and NPR.
Each area is represented by a square of the same dimensions.
<!--
Figure 3<!--\ref{fig:ggcartograms}-->d shows a hexagon tilemap representing each state using a single hexagon.

Recommended criteria to contrast mapping methods include average cartographic error, and maximum cartographic error, correct adjacency, maximum aspect ratio, and suitable relative positions [@ORC]. However, this does not consider the issues with actually producing rectangular cartograms. Algorithms for the creation of rectangular cartograms
-->

## Tile Map
<!-- 173 words -->
A tile map provides a tessellated display of consistent shapes. Each tile is usually one unit of measurement, this could be geographic regions such as states or population-based that use a consistent measure of population for each tile. Regions with over four neighbors require some necessary displacement. The tile map uses color to represent a value of a statistic for each area. A similar method to a rectangular cartogram represents each geographic area using a square of the same dimensions. The squares tessellate to create a grid. There are online media sources using this method, such as FiveThirtyEight, Bloomberg Business, The Guardian, The Washington Post, The New York Times and NPR. Tile maps may be difficult to create as they are best created manually, they require additional time and care as the number of geographic areas to include increases.
<!-- cite mosaic cartograms -->

Cano and others [@MDAC] define the term 'mosaic cartograms' for hexagonal tile displays, where the number of tiles for each area or the color of them can communicate the statistic of regions. 
When using several tile per regions, map makers can adjust the complexity of the boundaries in the resulting display. They can also make a trade-off between boundary complexity and simplicity by the size of the tiles used.

## Geofacet
<!-- 166 words -->
Hafen [@IGF] formalizes the term geofaceting to describe a grid display, the arrangement of tiles to create a grid that mimics the geographic topology of the set of areas. Like tile maps, they can reuse the arrangement of the areas for other spatial distribution visualizations. 
Geofaceting has the functionality of facets, often used to replicate visualizations for each subset of the data. Here, the data subsets are geographic. A tile map communicates one value per region in a visualization, geofaceting is a more flexible visualization for communication as it increases the amount of information displayed. Virtually any plot display can be shown in the tile representing the areas, allowing displays of multiple variables or values per geographic entity. Creating a tile map or a geofacet both require manual creation of the grid, except for the grids that other have already contributed.

```{r ggtilemap, fig.cap = "Two alternative displays, tile map (left) and geofaceted map (right), showing state age-adjusted rate of incidence for lung and bronchus in the USA. In the tile map, the layout approximates spatial location, with each state being an equal box filled with color representing cancer incidence. The geo-faceted map shows bar charts laid out in a grid approximating the spatial location of the state. The maps show age-adjusted rates for males and females. This display allows the presentation of multiple variables for each geographic area.", out.width="100%"}
knitr::include_graphics("figures/gggrids.png")
```

## Multivariate displays
<!-- 222 words -->

Pickle and others [@MMST] present linked micro map plots to match geographic and statistical data visually,  this serves as a solution to  multi-dimensionality issues. An online atlas of US states uses this display. These maps group areas based on their value for one variable, and additional columns provide displays that contrast the areas in each group by other variables.
The display juxtaposes choropleth maps and statistical plots;
it shows one map per group of the key separating variable, in a row with each additional statistical plot.
Linked micro maps predominantly use the choropleth map for displays of spatial relationships. These maps show spatial relationships by allotting spatial neighbors to the same group.
It is one of several alternative displays that allow maps to become bivariate displays, commonly used to present both an estimate and the associated uncertainty.

Bivariate choropleth maps [@VUADBC] blend color schemes to convey the intersection of categorized levels of an estimate and the associated uncertainty for each spatial area.
Lucchesi and Wikle [@VUADBC] also suggest map pixilation, it breaks each region into small pixels, and  allocates values to the individual pixels to create texture. This reflects the uncertainty around the area's estimate by randomly sampling from the confidence interval of the estimate of the area.
Animating these displays involves resampling the pixels for each frame.
Areas with uncertain values will flicker more dramatically than areas with more certain values.

# Comparison and critique of alternative displays {#sec:compare}
<!-- 408 words -->

```{r auscartograms, fig.cap = "Cartograms showing melanoma incidence in Australia: (a) contiguous, partially population transformed, (b) non-contiguous shape preserved, (c) Dorling, (d) hexagon tile map. The contiguous cartogram has expanded the highly populated areas while preserving the full shapes of rural areas. If it accurately sized areas by population, the country would be unrecognizable. The shape-preserved is unreadable due to the small area sizes. The Dorling cartogram presents all areas but many are difficult to compare. The hexagon tile map provides a reasonable spatial distribution despite having isolated hexagons in the outback areas.", message=FALSE, warning=FALSE, out.width='100%'}

aus1 <- png::readPNG("figures/aus_ggcont.png")
aus2 <- png::readPNG("figures/aus_ggncont.png")
aus3 <- png::readPNG("figures/aus_ggdorl.png")
aus4 <- png::readPNG("figures/aus_gghexmap.png")
load("aus_legend.rda")

load("figures/aus_legend.rda")

library(cowplot)
ggdraw() +
  draw_plot(rasterGrob(aus1), 0, .55, 0.5, 0.50) +
  draw_plot(rasterGrob(aus2), 0.5, 0.55, 0.5, 0.50) +
  draw_plot(rasterGrob(aus3), 0.0, 0.1, 0.5, 0.55) +
  draw_plot(rasterGrob(aus4), 0.5, 0.1, 0.5, 0.55) +
  draw_plot(aus_legend, 0, 0, 1, .1) +
  draw_plot_label(c("a", "b", "c", "d"), 
    c(0, .5, 0, 0.5), 
    c(1, 1, .5, .5), size = 20)
```

## Neither choropleth maps or cartograms perform well for Australia

Figure \ref{fig:auscartograms} shows four main types of cartograms displaying melanoma incidence on SA3 areas in Australia. The version of a contiguous cartogram (a) has expanded the highly populated areas while preserving the full shapes of rural areas. It has not fully resolved the population transformation of areas, and if it had accurately sized areas by population, the country would be unrecognizable. The shape-preserved cartogram is unreadable, and it has reduced all areas to tiny spots on the map. Zooming in on a high-resolution output shows it does preserve the areas. The Dorling cartogram has lost all geographic context. The hexagon tile map provides a reasonable spatial distribution despite having too much white-space in the outback areas.

## Pros and cons of different choices

Cartograms provide the spatial distortion to more accurately convey the statistical distribution, focuses on the human impact of the disease. However, if the population density is highly dissonant with geographic density, the cartogram will lose all spatial context, and thus futile. 
Dorling [@ACTUC] has a cartogram showing the 1966 general election results, which looked very little like the geographical shape of Australia.

Some mix of tiling, faceting or even micro maps, which allow some spatial continuity while also zooming into small areas, are good solutions for difficult geographies. Table \ref{tab:methods} summarizes the key criteria for testing maps and alternative displays. Moore and Carpenter [@SAMGIS] and Bell et al. [@CPISACA] provide suggestions and comments to help map creators best communicate their health data and spatial analysis.

<!-- choropleth maps, true topology, geographic distance, good for showing geographic importance 
To choose an appropriate map display, the map creator must consider the intended user, and message the map will communicate.
Choropleth displays utilize more traditional cartographic methods, they are usually true to the topography, displaying familiar boundaries of countries, states or administrative areas. Highlighting the geographic distribution, and differences between equal spatial units.
When there is a relationship between the variable being mapped and geographical features, a choropleth map would pair this information visually.
Choropleth methods require a decision to be made about the projection of the display, in this case aim to find a map projection that gives the least distorted representation of the geography [@UMPT].
However a choropleth is limited to a uni-variate display, presenting a spatial distribution by filling each area with a color.
These displays favor geographically large regions over small regions, though inset displays are often used to zoom in on a geographic area containing many small regions, this is seen in Figure 1<!--\ref{fig:choropleth-grid-create}--> c of the state of Queensland, where the capital city Brisbane is highlighted by the inset map. 
Users can identify the areas relevant to them by the familiar boundaries, and can easily contrast the experience of geographic neighbors. The map is reusable and fit for many purposes.
-->

<!-- Cartograms, sacrifice geography, focus on the outliers/ relationships in data space 
Cartographers were historically the creators of geographic displays, and epidemiologists utilized the maps that had been prepared in advance.
Cartograms with the population variable used to distortion the geographic map regions and boundaries are density equalizing maps, that reduce the visual prominence of large, but low-population areas [@VSSDCUC]. 
Cartogram displays incorporate the statistics and population into the design of the display,
shifting and sometimes sacrificing familiar boundaries to draw attention to large outliers in the data space or population density.
The unusual shapes are not longer familiar and may be difficult to compare. This can prevent effective comparison of the statistic. Cartograms also favor extreme statistics in the positive direction. Small extreme values are often lost due to the resulting small sized areas the do not draw user's attention.
The change in the map based on one variable provides the opportunity for bi-variate displays, using the change in area for one variable, and color for another.
The difference between the familiar map display helps highlight the impact of the disease on communities.
-->

<!-- Alternatives, remove geographic shapes, keep spatial relationship, focus on the outliers/ relationships in population space 
Alternative maps may replace the unexpected boundaries and shapes with familiar, simple shapes. This makes the spatial distribution the primary concept communicated by the display.
Areas that are geographically close will maintain connectedness in some way, but the population and the statistic will dominate.
tilemap displays may only allow one variable to be visualized if each region is given consistent map space, this occurs in geofaceting. 
Dorlings provide bi-variate displays as they use the size of the shape for one variable, and color for another.
These alternatives are especially helpful for data aggregations where administrative boundaries break populations into groups.
The experience of each community may be worth considering, as the experiences shared by the population within them may be similar due to the services and facilities they share.
This display allows a more equitable view of each community, and does not minimise those that operate on a smaller geographic scale.

Faceting provides an opportunity to contrast the overall distribution of variables across a geographic space or collection of geographic areas.
This display makes is difficult to compare specific regions.-->

Table: \label{tab:methods} Summary of features and constraints of common mapping methods used to display cancer statistics (Y=Yes, N=No, S=Sometimes).

|                     | Choropleth | Contiguous | Non-contig | Dorling | tile maps | Geofacets |
|:-----------|:------------:|:-----------:|:---------------:|:---------:|:-----------:|:----------:|
|Spatial distortion   |  N  |  Y  |  Y  |  Y  |  Y  |  Y  |
|Preserves neighbors  |  Y  |  Y  |  Y  |  S  |  S  |  S  |
|Conceals small areas |  Y  |  S  |  N  |  N  |  N  |  N  |
|Uniform shape        |  N  |  N  |  N  |  Y  |  Y  |  Y  |
|Univariate only      |  Y  |  Y  |  Y  |  S  |  S  |  N  |
|Manual construction  |  N  |  N  |  N  |  N  |  Y  |  Y  |

<!-- Creating maps of diseases now involves more decisions to be made by map makers, rather than cartographers. 
Technology has played an enormous role in increasing the opportunities for map makers.
The computation and graphics power have made creation, alteration and interactivity possible.
as these options have expanded and it is the objectives of the investigator that will drive the choices. @CPISACA and @SAMGIS have provided suggestions and comments to help map creators best communicate their health data and spatial analyses.-->

# User interaction {#ch:interacting}
<!-- 634 words -->

<!-- Interactivity and animation can add value to online atlases, by enabling additional information to be presented on-the-fly with one display. It also invites the viewer to engage with the display, and the information.

The word "interactive" is used, and perhaps, confused for many different graphics actions. At one extreme, it is synonymous with direct manipulation of the display, such as selecting an area, mousing over, or panning and zooming, using mouse action on the plot. At the other extreme, it refers to graphical user interfaces (GUI) such as selecting an item on a menu to change the variable being displayed, or checking a toggle box, or dragging a scrollbar. @RPIG provides a good overview of types of interactivity with plots. a key difference from an animation is that interactivity typically involves on-the-fly computations, and changes to the plot.

The primary interactive tools found in online cancer maps include

- drop-down menus to select features,
- variable selection, 
- tool tips, 
- zooming and panning.

The latter two would be considered to be direct manipulation on the map display, and the former two are GUI interfaces. The effects of these actions are rapid, incremental, and reversible [@DMIV]. 
-->

One of the concerns of adding too much information to a map is the fear of cognitive overload [@mcgranaghan1993cartographic] in which the user reaches an information threshold, beyond which will not be able to make sense of the information. It can be a juggling act for a diverse audience, with experts probably preferring more detail [@cliburn2002design]. Interactivity is a design feature within modern mapping methods that can be used to incorporate additional information and complexity without overloading the user. Effective  user-centred interactive actions produce rapid, incremental, and reversible changes to the display [@DMIV].

Monmonier [@HTLWM] says that interactivity can be used to allow users to explore the map for more information and provides flexibility for the display. The user can toggle between different variables, map views or even multiple realizations of future scenarios [@goodchild1994introduction]. This provides additional mechanisms for the users to digest the uncertainty of the available information [@maceachren1992visualizing, @van1994visualization]. When the needs of the audience are changeable and are also the priority, the map creator can allow interactivity for map users to explore a data set through dynamic interactions. This can allow inspection of the data from many views [@DQBCM]. User interaction with maps helps to understand and interpret the spatial distribution of disease, to validate, explain or explore the presented statistics and their relationships to each other [@TNTEA].

Interactivity enables supplementary information to be incorporated into online atlases without cluttering the display. Interactive design features, found in online cancer maps, include tool tips, drop-down menus, data selection, zooming, and panning allow users to explore the map as they want more information and allow flexibility in the display [@HTLWM]. The use of these supports can be found in various online cancer maps [@roberts2019communication].

Animation, in contrast to interactivity, usually involves pre-computing views and showing these in a sequence. An overview of animation in association with the R package `gganimate` [@gganimate], with some focus on maps, explaining transitions between frames in a linear, to provide a fixed narrative to passive users [@TGA]. Animations are used to communicate a message by capturing and directing users' attention. It is most often employed to show changes over time. The controls for basic animation are usually placed outside of the plot space [@TGA], the map image is updated/replaced as the animation progresses.

Weather maps are a thoroughly developed example of animation of spatial displays to communicate information to the general public [@CPISACA]. The movement of a weather system will follow a forecasted path, all map users can follow the animated path of the weather system across the geography over a specified period.

The Australian Cancer Atlas [@TACA] provides [tours](
https://atlas.cancer.org.au/app/tour/lungcancer) that change the display to draw users' attention to areas on the map that are relevant to the story.
This implementation of animation gives users tools to plan their exploration.

<!-- 
XXX This part needs to go to interactivity!
One of the concerns of adding too much information to a map is the fear of cognitive overload [@mcgranaghan1993cartographic] in which the user reaches an information threshold, beyond which will not be able to make sense of the information. These concerns are not unfounded, complexity and density of representation methods appear to overwhelm novice decision makers, while experts are able to use the detail more readily when making-decisions [@cliburn2002design]. Interactivity is a design feature within modern mapping methods that can be used to incorporate additional information and complexity without overloading the user.

Interactivity enables supplementary information to be incorporated into online atlases without cluttering the display. Interactive design features, found in online cancer maps, include tooltips, drop-down menus, data selection, zooming and panning allow users to explore the map as they want more information and allow flexibility in the display [@HTLWM]. The use of these supports were found in various online cancer maps identified by @roberts2019communication. The controls for basic interactive features are often placed outside of the plot space [@TGA], thus the map image is updated/replaced as the user interacts with the controls. For example, changing the population age or other demographic variables. Some more advanced interactions include direct interactions with the plot via the use of overlaid tool tip features, very few cancer atlases involve these more complex selection tools.

Additionally, interactivity allows the user to toggle between different variables, map views or multiple realizations of possible future scenarios [@goodchild1994introduction]. Thus providing additional mechanisms for the users comprehension as well as the uncertainty of the available information (@maceachren1992visualizing; @van1994visualization).

These interactive features provide an opportunity for users to explore the additional information available. This helps users to understand and interpret the spatial distribution presented, as well as validate, explain or explore the presented statistics and their relationships to each other and/or their underlying spatial distribution. This allows relationships between spatial areas and diseases to be explored with sophistication in non-traditional but still 'cognitively accessible' ways [@TNTEA]. The interactive features of the publicly available maps identified by @roberts2019communication allow the exploration of geographic hierarchies, population distribution, statistical uncertainty, demographics and socio-economic indicators.
@TNTEA suggested LM plots as a solution to linking cartography and statistical graphics.
-->

## Interaction and animation in publicly available atlases

Many publicly available cancer atlases employ interactivity and/or animation. Figure 6 <!--\ref{fig:interacting}--> shows the interactive GUI tools provided with the public atlases. Mostly these are menus, and toggles, for choosing indicators, and subsets of the population.

## Under-utilized techniques

Figure 7 <!--\ref{fig:animating}--> shows two examples of more sophisticated interactive maps. The Spanish Cancer map (left) contains a linked display between a choropleth map and time series plots of cancer change. In linked plots, changing values in one display will trigger changes of corresponding elements in another display. Here, the temporal change in the choropleth map can be played out as an animation. Mousing over the time series plots will highlight the line for a particular region. The Canadian Breast Cancer Mortality map (right) has a magnifying glass that allows the user to zoom into small areas. It is easy to control and shows precise details in small areas.

<!-- It is uncommon to see linked displays utilized in online atlases, but there are several frameworks that support interactive displays of multivariate data. @TNTEA suggested LM plots as a solution to linking cartography and statistical graphics. 
Linked displays allow for direct manipulation, @RPIG uses Swayne and Klinke's definition of high level interaction that occurs via data selection or brushing tools that highlight subsets of graphical elements such as polygons or points. Selections can be made statistically through interactions which allow subsets to be selected from statistical displays, such as tables of data.
Monmonier implemented linking choropleth maps with statistical plots, adding a geographic component to scatter plots [@MCRG], the use of link allows spatial dimensions to be reserved for geographic locations of areas. The statistical variation of additional variables can be presented in other appropriate displays.
Low level interaction is more common, this occurs via the user interface or the map.
Interacting with the map returns values or information used in the graphical display without referring to, of changing, the original data source of the map.
-->

<!-- Animations -->
<!--Animations communicate a story, showcasing a critical region by defining and progressing through a small set of keyframes, or waypoints [@ESRI].
@TGA suggests the use of moving objects demands attention by taking advantage of the behaviour of the visual cognition system.
They can be produced to not tolerate or require user input, creating useful, directed descriptions of processes, such as highlighting a particular region or time period [@HTLWM].
Animations ensure all users follow the same interpretive path but creators cannot control the level of engagement.
-->

<!-- Animations can be implemented to allow users varying degrees of control.
Figure \ref{fig:animating} a gives an example of how time can be included through animating displays. This is a basic form of interactivity with animation allowing users to play, pause and rewind films, where the film is intended to be watched in one direction but the position in the animation can be controlled.
@CPISACA provide weather maps as a thoroughly developed example of animation of spatial displays to communicate information to the general public.
The movement of a weather system will follow a forecasted path, all map users can follow the animated path of the weather system across the geography over a specified period of time.
Where the message is most important, static or animated graphics allow control over the display and interpretation.
User interaction more allow clicks to progress between stages of more complex animations.
The Australian Cancer Atlas [@TACA] provides [tours](
https://atlas.cancer.org.au/app/tour/lungcancer) which change the display to draw user's attention to areas on the map that are relevant to the story.
This implementation of animation gives users tools to plan their own exploration.
-->

```{r interacting, fig.cap = "Interactive controls of displays in publicly available choropleth cancer maps: (i) GUI controls for statistic, sex, age groups, continents, and cancer types for [Globocan 2018: Cancer Today ](http://globocan.iarc.fr/Pages/Map.aspx), (ii) Menus for variable selection and zooming on [Bowel Cancer Australia Atlas](http://www.bowelcanceratlas.org/), (iii) Menus for choosing variables and countries in [The Cancer Atlas ](http://canceratlas.cancer.org/data/#?view=map&metric=INCID_ALL_M),  (iv) Tabs for different indicators and cancer types in [Global Cancer Map](http://globalcancermap.com/), (v) Menus and toggles for variable and subset selection in [United States Cancer Statistics: Data Visualizations](https://gis.cdc.gov/Cancer/USCS/DataViz.html).", fig.width = 12, fig.height = 9}
i1 <- png::readPNG('cancer_map_images/interactivity/3.interactivity.png') # b in choro grid
i2 <- png::readPNG('cancer_map_images/interactivity/15.interactivity.png') # d in choro grid
i3 <- png::readPNG('cancer_map_images/interactivity/4.interactivity.png')
i4 <- png::readPNG('cancer_map_images/interactivity/5.interactivity.png')
i5 <- png::readPNG('cancer_map_images/interactivity/7.interactivity.png') # e in choro grid

ggdraw() +
  draw_plot(rasterGrob(i1), 0,  0.13,   .28,  .83) + #a
  draw_plot(rasterGrob(i2), 0.32, 0.56, .68, .4) + #b
  draw_plot(rasterGrob(i3), 0.32, 0.41, .68, .13) + #c
  draw_plot(rasterGrob(i4), 0.32, 0.14, .68, .25) + #d
  draw_plot(rasterGrob(i5), 0, 0, 1, .13) + #e
  draw_plot_label(c("i", "ii", "iii", "iv", "v"), 
    c(0.01, 0.3, 0.3, 0.3, 0.01), 
    c(0.99, 0.99, 0.57, 0.40, 0.15), size = 20)

ggsave(filename = "figures/supp_grid.png", device = "png", 
  width = 12, height = 9, dpi = 300)
```

```{r animating, fig.cap = "Two examples of advanced interactivity (and animation) in publicly available choropleth cancer maps: a. Linked maps and time-series line plots, with temporal animation in [Map of Cancer Mortality Rates in Spain](http://www.geeitema.org/AtlasET/atlas.jsp?causa=e27MEPOC), b. A highly responsive magnifying glass on a map of [Breast Cancer Mortality in Canada](http://www.ehatlas.ca/light-pollution/maps/breast-cancer-mortality).", fig.width = 12, fig.height = 9}
a1 <- png::readPNG("cancer_map_images/animation/23.animation.png")
a2 <- png::readPNG("cancer_map_images/animation/2.animation.png")

ggdraw() +
  draw_plot(rasterGrob(a1), 0, 0, 0.3, 1) +
  draw_plot(rasterGrob(a2), 0.3, 0, 0.7, 1) +
  draw_plot_label(c("I", "II"), 
    c(0, .3), c(1, 1), size = 20)
```

# Conclusions
<!-- 307 words -->

<!-- summary of mapping -->
This paper provides an overview of mapping practices as used for cancer atlases and new approaches that could be adopted. The conventional approach is the choropleth map, and it is widely used. When there are small areas, as occurs in Australia where the population is concentrated on the coast, the information about cancer can be lost, and alternatives are needed. Making an inset can clarify congested regions but this breaks the viewers' attention because they need to shift focus from the map to the inset, and if there are many congested areas, many insets would be needed. The map alternatives, like cartograms and their variations, can be useful to allow the spatial distribution of cancer data to be digested.

<!-- statistics -->
Many different statistics are commonly used for cancer displays. The most basic is the incidence rate, which is easy to understand. It is common to see relative rates in many maps which measure how far a region is above or below the average. The purpose of using a relative rate is, perhaps the desire to pinpoint the areas that need attention because they have higher than expected rates. However, we lose the incidence rate information and thus interpretability. A region might be much higher than average, but it may not be close to a health concern, because all regions have a low incidence.

<!-- interaction -->
Interaction with maps is an important component of public atlases. A purpose is to provide access to more information than is possible to display in a single map, without overwhelming the viewer. Too many choices though can similarly overwhelm a viewer, and thus decisions do need to be made about content to provide, or hide, for accurate and comprehensive communication of information. Similarly, providing ways for users to interact with the display encourages engagement, and creative, efficient, elegant, interactive tools elicit curiosity about the data.

# Acknowledgements
<!-- 77 words -->
The authors would like to thank Dr. Earl Duncan, Professor Kerrie Mengersen, Dr. Susanna Cramb and Dr. Peter Baade for conversations on the content of this article.

The following R [@R] packages were used to produce this paper: tidyverse [@tidyverse], RColorBrewer [@RColorBrewer], ggthemes [@ggthemes], png [@png], cowplot [@cowplot], sf [@sf], spData [@spData], cartogram [@cartogram], sugarbag [@sugarbag], knitr [@knitr], rmarkdown [@rmarkdown].

Files to reproduce the paper, and code to reproduce the plots, are available at https://github.com/srkobakian/review.

# References

<!--
## counts versus predictions

@SEMA
'Disease mapping is carried out to summarise spatial and spatio-temporal variation in risk.'
Descriptive statistic of the relationship between geography and a disease.
population, exposure, and health data may have point data or aggregated summaries (count data)
potentially be subject to a number of inaccuracies
diagnostic error or misclassification

Not only the counts, but population, migration, deaths and births are modelled across census years.

rapid visual summary of complex geographical information, descriptive purposes, to generate hypotheses, for surveillance, and to aid policy formation and resource allocation.

When using counts:
 small populations imply large variability in estimated rates

Decisions taken at this stage may be critical as apparent ‘clusters’ may depend crucially on the boundaries chosen in time or space: ‘The more narrowly the underlying population is defined, the less will be the number of expected cases, the greater will be the estimate of the excess rate, and often the more pronounced will be the statistical significance’

Using prediction:
assuming that neighboring μi are correlated leads to a ‘clustering’ prior

-->

<!-- presents forms of low level interactions with a user interface, these are the control menus for several online cancer atlases.
The user interface for these online atlases allow combinations of cancer type, gender, and other factors to be shown on the map display and the user to move between and/or compare different views of these variables. 
These controls may also change the geographies shown, such as continents or countries, states or counties. These displays can show familiar shapes for users to orient themselves on the map, then zoom in to explore densely populated communities.
Interactions with online atlases can also change the amount of detail shown.-->

<!-- Interactivity can be implemented within modern mapping methods to incorporate additional information and complexity without overloading the user.
Adding too much information to a map may create cognitive overload [@mcgranaghan1993cartographic] in which the user reaches an information threshold, beyond which will not be able to make sense of the information. 
Complexity and density of visualization methods can overwhelm novice decision makers however, experts are able to use the detail more readily when making-decisions [@cliburn2002design]. 
Supplementary information can be incorporated into online atlases without cluttering the display. Interactive design features found in online cancer maps include tool tip features, drop-down menus, data selection, zooming and panning.
Allowing users to explore the map for more information and providing flexibility in the display [@HTLWM]. The use of these supports were found in various online cancer maps identified by @roberts2019communication.
Additionally, interactivity allows the user to toggle between different variables, map views or multiple realizations of possible future scenarios [@goodchild1994introduction]. Thus providing additional mechanisms for the users comprehension as well as the uncertainty of the available information [@maceachren1992visualizing, @van1994visualization]. For example, changing the population, age or other demographic variables.
The controls for basic interactive features are often placed outside of the plot space [@TGA], the map image is updated/replaced as the user interacts with the controls. 
Some more advanced interactions include direct interactions with the plot via the use of overlaid tool tip features, very few cancer atlases involve these more complex selection interactive capabilities.

Relationships between spatial areas and diseases can be explored with sophistication in nontraditional but still 'cognitively accessible' ways [@TNTEA]. This helps users to understand and interpret the spatial distribution presented, as well as validate, explain or explore the presented statistics and their relationships to each other and/or their underlying spatial distribution.

The interactive features of the publicly available maps identified by @roberts2019communication allow the exploration of geographic hierarchies, population distribution, statistical uncertainty, demographics and socio-economic indicators.

Interactivity and animation provide an alternative to the traditional approach of a single map for a single use, and enable the map to become more exploratory.
Interactivity allows users to drive their own exploration, controlling a display through variable controls, links, tooltips, and data selection tools [@TGA]. tooltips allow users to obtain additional information regarding specific geographic areas.

Development and adoption of this technology has allowed map users more control of the displayed message. 
-->

<!--to expand the communicative ability of a disease map. It is important to identify the primary questions that might be asked by the users of a map, and allow the display to change to best answer these questions. @CPISACA provide three questions that are generally asked when using a disease map:

- What is the mortality rate in a certain area?
- Are there geographic trends in the data, or regions of unusually high or low rates?
- Is the lung cancer mortality pattern similar to the pattern of smoking prevalence shown in a companion map?

A single map display may not effectively answer all three at the same time.
This is especially true when some areas are small and difficult to identify clearly in a single display or resolution. However, with the ability to interact with a map, or animate the display, map creators can specify and answer these questions by directing the attention of a map user to the areas of interest.
-->